---
title: "Binomial Proportion model"
author: "David Kent (dk657)"
date: "January 9, 2018"
output:
  md_document:
    toc: true
---

```{r echo=F, message=F, warning=F}
options(scipen=100,digits=2)
library(glmmTMB)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
#library(nlme) # check that estimates match
#library(GAMM) # check that estimates match
library(knitr)
library(kableExtra)
```

Read in data:

```{r, warning=F}
san <- read.csv("SummarySanitizer_2018705.csv") %>%
  mutate(Trial=as.factor(Trial),
         PAA=PAA*10.1/10,
         prop=Control/PAA,
         reduc=log10(Control/PAA),
         PAA_v=.1/10.1/10^DilutionP) %>%
  mutate(PAA_detect=1/PAA_v) %>%
  mutate(sanity=ifelse(PAA==0,TRUE,PAA>=PAA_detect),
         right_censored=ifelse(prop==Inf,"censored","not_censored"),
         right_censor_limit_prop=Control/PAA_detect,
         right_censor_limit_reduc=log10(Control/PAA_detect),
         PAA=round(PAA,digits = 0))
```

# Listeria model

```{r}
l_data <- san %>%
  filter(Strain!="FSL C2-0008") %>% # remove surrogate?
  filter(PathogenGroup=="Listeria monocytogenes")
l_data_w_surrogate <- san %>%
  filter(PathogenGroup=="Listeria monocytogenes")
```

## Fit binomial proportion model

We'll fit a crossed random-effects log-link binomial proportion model (with OLRE
to correct for overdispersion) to the response
$\log(N_{\text{san}}/N_{\text{cont}})$, the negative of which is our
"log-reduction."

```{r warning=F}
l_data$olre <- factor(row(l_data)[,1])
l_bin <- glmer(PAA/Control~(1|Strain)+(1|Condition)+(1|Strain:Condition)+(1|olre),
          weights=Control,
          family=binomial(link="log"),
          control=glmerControl(optimizer="Nelder_Mead"),
          data=l_data)

# Extract BLUPs and intercept from model
l_strain_effs <- ranef(l_bin)$Strain %>% tibble::rownames_to_column(var="Strain")
l_condition_effs <- ranef(l_bin)$Condition %>% tibble::rownames_to_column(var="Condition")
l_int_effs <- ranef(l_bin)$`Strain:Condition` %>% tibble::rownames_to_column(var="Int")
l_intercept <- fixef(l_bin)[1]

# Compute BLUP of "log-reduction"
l_blups <- l_data %>%
  select(Strain,Condition) %>%
  mutate(Int=paste(Strain,Condition,sep=":")) %>%
  distinct() %>%
  left_join(l_strain_effs,by="Strain") %>% rename(strain_eff=`(Intercept)`) %>%
  left_join(l_condition_effs,by="Condition") %>% rename(condition_eff=`(Intercept)`) %>%
  left_join(l_int_effs,by="Int") %>% rename(int_eff=`(Intercept)`) %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         Int=factor(Int)) %>%
  mutate(intercept=l_intercept) %>%
  mutate(BLUP=log10(exp(1))*(-intercept-strain_eff-condition_eff-int_eff))

# Add zeroes for plotting
l_blups <- l_data %>% 
  group_by(Strain,Condition) %>%
  summarize(zeroes=sum(PAA==0)) %>%
  left_join(l_blups,by=c("Strain","Condition")) %>%
  ungroup()
```

Giving this model:
```{r}
summary(l_bin)
```

With the following CIs for variance components:
```{r warning=F}
#confint(l_bin) %>%
#  kable()
```

This gives us the following BLUPs:
```{r}
l_strain_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()

l_condition_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()

l_int_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()
```

## Plotting the results

```{r}
l_blups %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,BLUP,zeroes) %>%
ggplot(aes(x=i_Strain,y=BLUP,color=Condition,linetype=Condition)) +
  geom_line() +
  geom_text(aes(label=zeroes)) +
  ylim(0,6.5) +
  labs(x="Strain",y="Binomial proportion model",title=expression(italic("Listeria"))) +
  scale_x_continuous(breaks=1:length(levels(factor(l_blups$Strain))),
                     labels=levels(factor(l_blups$Strain))) +
  theme_bw()
```

Compare to the mean estimates (with detection limit substituted where sanitizer
count is zero)

```{r}
l_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Strain,y=mean,color=Condition,linetype=Condition)) +
  geom_line() +
  geom_point() +
  labs(x="Strain",y="Mean",title="Listeria") +
  ylim(0,6.5) +
  scale_x_continuous(breaks=1:length(levels(factor(l_data_w_surrogate$Strain))),
                     labels=levels(factor(l_data_w_surrogate$Strain))) +
  theme_bw()
```

Looking at this the other way:

```{r}
l_blups %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,BLUP,zeroes) %>%
ggplot(aes(x=i_Condition,y=BLUP,color=Strain,linetype=Strain)) +
  geom_line() +
  geom_text(aes(label=zeroes)) +
  ylim(0,6.5) +
  labs(x="Condition",y="Binomial proportion model",title="Listeria") +
  scale_x_continuous(breaks=1:length(levels(factor(l_blups$Condition))),
                     labels=levels(factor(l_blups$Condition))) +
  theme_bw()
```

Compare to the mean estimates (with detection limit substituted where sanitizer
count is zero)

```{r}

l_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Condition,y=mean,linetype=Strain, pch=Strain)) +
  geom_line(size=1) +
    geom_point(size=2.5) +
  labs(x="Condition",y="log reduction CFU/mL",title=expression(italic("Listeria")))+
  ylim(0,9) +
  scale_x_continuous(breaks=1:length(levels(factor(l_data_w_surrogate$Condition))),
                     labels=levels(factor(l_data_w_surrogate$Condition))) +
  theme_bw(base_size = 20)

```


## Prediction intervals

```{r}
# This seems more direct, but I'm not sure how to handle conditioning on a random effect
#l_newdat <- unique(l_data[,c("Strain","Condition")])
#mm<-model.matrix(~1,l_newdat)
#y<-mm%*%fixef(l_bin)
#pvar1 <- diag(mm %*% tcrossprod(vcov(l_bin),mm))
#tvar1 <- pvar1 + VarCorr(l_bin)$olre[1] + VarCorr(l_bin)$`Strain:Condition`[1] +
#  VarCorr(l_bin)$Condition[1] + VarCorr(l_bin)$Strain[1]
#l_newdat <- data.frame(
#  y=exp(y),
#  plo = -log10(exp(y-1.96*sqrt(pvar1))),
#  phi = -log10(exp(y+1.96*sqrt(pvar1))),
#  tlo = -log10(exp(y-1.96*sqrt(tvar1))),
#  thi = -log10(exp(y+1.96*sqrt(tvar1)))
#)
```

```{r}
mu <- l_bin@beta[1]
sds <- l_bin@theta
l_lin_pred_overall <- function(x) {
  -log10(exp(mu + 
        rnorm(1,0,sds[1]) +
        rnorm(1,0,sds[2]) +
        rnorm(1,0,sds[3]) +
        rnorm(1,0,sds[4])))
}
l_overall_sim <- data.frame(fixed_feature="None",
           fixed="None",
           sim=unlist(lapply(1:1e4,FUN=l_lin_pred_overall)))

l_lin_pred_strain_fixed <- function(x,strain) {
  -log10(exp(mu + 
        rnorm(1,0,sds[1]) +
        rnorm(1,0,sds[2]) +
        rnorm(1,0,sds[3]) +
        strain))
}
l_strain_sim <- data.frame(fixed_feature=NULL,
                     fixed=NULL,
                     sim=NULL)
for( s in l_strain_effs$Strain ) {
  tmp <- data.frame(fixed_feature="Strain",
                    fixed=s,
                    sim=unlist(
                      lapply(1:1e4,
                             FUN=l_lin_pred_strain_fixed,
                             strain=l_strain_effs$`(Intercept)`[l_strain_effs$Strain==s])))
  l_strain_sim <- rbind(l_strain_sim,tmp)
}

l_lin_pred_condition_fixed <- function(x,condition) {
  -log10(exp(mu + 
               rnorm(1,0,sds[1]) +
               rnorm(1,0,sds[2]) +
               condition +
               rnorm(1,0,sds[4])))
}
l_condition_sim <- data.frame(fixed_feature=NULL,
                     fixed=NULL,
                     sim=NULL)
for( s in l_condition_effs$Condition ) {
  tmp <- data.frame(fixed_feature="Condition",
                    fixed=s,
                    sim=unlist(
                      lapply(1:1e4,
                             FUN=l_lin_pred_condition_fixed,
                             condition=l_condition_effs$`(Intercept)`[l_condition_effs$Condition==s])))
  l_condition_sim <- rbind(l_condition_sim,tmp)
}

l_pred <- rbind(l_overall_sim,
                l_strain_sim,
                l_condition_sim)
l_pred %>%
  group_by(fixed,fixed_feature) %>%
  summarize(mean=mean(sim),
            lower=quantile(sim,0.025),
            upper=quantile(sim,0.975)) %>%
ggplot(aes(x=fixed,y=mean)) +
  geom_point() +
 labs(x="fixed effects",y="Log reduction CFU/mL",title=expression(italic("L. monocytogenes"))) +
  geom_errorbar(aes(ymin=lower,ymax=upper),width=0.2) +
  facet_grid(~fixed_feature,scales = "free") +
  theme(axis.text.x=element_text(angle = 45,hjust=1))
```

# Salmonella model

```{r}
s_data <- san %>%
  filter(Strain!="FSL R9-5275") %>% # remove surrogate?
  filter(Strain!="FSL R9-6232") %>% # remove surrogate?
  filter(PathogenGroup=="Salmonella")
# Substitute control count where PAA count was higher
s_data$PAA <- pmin(s_data$PAA,s_data$Control-1)
s_data_w_surrogate <- san %>%
  filter(PathogenGroup=="Salmonella")
```

## Fit binomial proportion model

We'll fit a crossed random-effects log-link binomial proportion model (with OLRE
to correct for overdispersion) to the response
$\log(N_{\text{san}}/N_{\text{cont}})$, the negative of which is our
"log-reduction."

```{r warning=F}
s_data$olre <- factor(row(s_data)[,1])
s_bin <- glmer(PAA/Control~(1|Strain)+(1|Condition)+(1|Strain:Condition)+(1|olre),
          weights=Control,
          family=binomial(link="log"),
          #control=glmerControl(optimizer="Nelder_Mead"),
          data=s_data)

# Extract BLUPs and intercept from model
s_strain_effs <- ranef(s_bin)$Strain %>% tibble::rownames_to_column(var="Strain")
s_condition_effs <- ranef(s_bin)$Condition %>% tibble::rownames_to_column(var="Condition")
s_int_effs <- ranef(s_bin)$`Strain:Condition` %>% tibble::rownames_to_column(var="Int")
s_intercept <- fixef(s_bin)[1]

# Compute BLUP of "log-reduction"
s_blups <- s_data %>%
  select(Strain,Condition) %>%
  mutate(Int=paste(Strain,Condition,sep=":")) %>%
  distinct() %>%
  left_join(s_strain_effs,by="Strain") %>% rename(strain_eff=`(Intercept)`) %>%
  left_join(s_condition_effs,by="Condition") %>% rename(condition_eff=`(Intercept)`) %>%
  left_join(s_int_effs,by="Int") %>% rename(int_eff=`(Intercept)`) %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         Int=factor(Int)) %>%
  mutate(intercept=s_intercept) %>%
  mutate(BLUP=log10(exp(1))*(-intercept-strain_eff-condition_eff-int_eff))

# Add zeroes for plotting
s_blups <- s_data %>% 
  group_by(Strain,Condition) %>%
  summarize(zeroes=sum(PAA==0)) %>%
  left_join(s_blups,by=c("Strain","Condition")) %>%
  ungroup()
```

Giving this model:
```{r}
summary(s_bin)
```

With the following CIs for variance components:
```{r warning=F}
confint(s_bin,method="boot",nsim=100) %>%
  kable()
```

This gives us the following BLUPs:
```{r}
s_strain_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()

s_condition_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()

s_int_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()
```

## Plotting the results

```{r}
s_blups %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,BLUP,zeroes) %>%
ggplot(aes(x=i_Strain,y=BLUP,color=Condition,linetype=Condition)) +
  geom_line() +
  geom_text(aes(label=zeroes)) +
  ylim(0,9) +
  labs(x="Strain",y="Binomial proportion model",title="Salmonella") +
  scale_x_continuous(breaks=1:length(levels(factor(s_blups$Strain))),
                     labels=levels(factor(s_blups$Strain))) +
  theme_bw()
```

Compare to the mean estimates (with detection limit substituted where sanitizer
count is zero)

```{r}
s_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Strain,y=mean,color=Condition,linetype=Condition)) +
  geom_line() +
  geom_point() +
  labs(x="Strain",y="Mean",title="Salmonella") +
  ylim(0,9) +
  scale_x_continuous(breaks=1:length(levels(factor(s_data$Strain))),
                     labels=levels(factor(s_data$Strain))) +
  theme_bw()
```

Looking at this the other way:

```{r}
s_blups %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,BLUP,zeroes) %>%
ggplot(aes(x=i_Condition,y=BLUP,color=Strain,linetype=Strain)) +
  geom_line() +
  geom_text(aes(label=zeroes)) +
  ylim(0,9) +
  labs(x="Condition",y="Binomial proportion model",title="Salmonella") +
  scale_x_continuous(breaks=1:length(levels(factor(s_blups$Condition))),
                     labels=levels(factor(s_blups$Condition))) +
  theme_bw()
```

Compare to the mean estimates (with detection limit substituted where sanitizer
count is zero)

```{r}
s_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Condition,y=mean,linetype=Strain, pch=Strain)) +
  geom_line(size=1) +
  geom_point(size=2.5) +
  labs(x="Condition",y="Log reduction CFU/mL",title=expression(italic("Salmonella"))) +
  ylim(0,9) +
  scale_x_continuous(breaks=1:length(levels(factor(s_data$Condition))),
                     labels=levels(factor(s_data$Condition))) +
  theme_bw(base_size = 20)
```

## Prediction intervals

```{r}
mu <- s_bin@beta[1]
sds <- s_bin@theta
s_lin_pred_overall <- function(x) {
  -log10(exp(mu + 
        rnorm(1,0,sds[1]) +
        rnorm(1,0,sds[2]) +
        rnorm(1,0,sds[3]) +
        rnorm(1,0,sds[4])))
}
s_overals_sim <- data.frame(fixed_feature="None",
           fixed="None",
           sim=unlist(lapply(1:1e4,FUN=s_lin_pred_overall)))

s_lin_pred_strain_fixed <- function(x,strain) {
  -log10(exp(mu + 
        rnorm(1,0,sds[1]) +
        rnorm(1,0,sds[2]) +
        rnorm(1,0,sds[3]) +
        strain))
}
s_strain_sim <- data.frame(fixed_feature=NULL,
                     fixed=NULL,
                     sim=NULL)
for( s in s_strain_effs$Strain ) {
  tmp <- data.frame(fixed_feature="Strain",
                    fixed=s,
                    sim=unlist(
                      lapply(1:1e4,
                             FUN=s_lin_pred_strain_fixed,
                             strain=s_strain_effs$`(Intercept)`[s_strain_effs$Strain==s])))
  s_strain_sim <- rbind(s_strain_sim,tmp)
}

s_lin_pred_condition_fixed <- function(x,condition) {
  -log10(exp(mu + 
               rnorm(1,0,sds[1]) +
               rnorm(1,0,sds[2]) +
               condition +
               rnorm(1,0,sds[4])))
}
s_condition_sim <- data.frame(fixed_feature=NULL,
                     fixed=NULL,
                     sim=NULL)
for( s in s_condition_effs$Condition ) {
  tmp <- data.frame(fixed_feature="Condition",
                    fixed=s,
                    sim=unlist(
                      lapply(1:1e4,
                             FUN=s_lin_pred_condition_fixed,
                             condition=s_condition_effs$`(Intercept)`[s_condition_effs$Condition==s])))
  s_condition_sim <- rbind(s_condition_sim,tmp)
}

s_pred <- rbind(s_overals_sim,
                s_strain_sim,
                s_condition_sim)
s_pred %>%
  group_by(fixed,fixed_feature) %>%
  summarize(mean=mean(sim),
            lower=quantile(sim,0.025),
            upper=quantile(sim,0.975)) %>%
ggplot(aes(x=fixed,y=mean)) +
  geom_point() +
 labs(x="fixed effects",y="Log reduction CFU/mL",title=expression(italic("Salmonella"))) +
  geom_errorbar(aes(ymin=lower,ymax=upper),width=0.2) +
  facet_grid(~fixed_feature,scales = "free") +
  theme(axis.text.x=element_text(angle = 45,hjust=1))
```

# E. coli model

```{r}
e_data <- san %>%
  filter(Strain!="FSL R9-4077") %>% # remove surrogate?
  filter(PathogenGroup=="Escherichia coli")
e_data_w_surrogate <- san %>%
  filter(PathogenGroup=="Escherichia coli")
```

## Fit binomial proportion model

We'll fit a crossed random-effects log-link binomial proportion model (with OLRE
to correct for overdispersion) to the response
$\log(N_{\text{san}}/N_{\text{cont}})$, the negative of which is our
"log-reduction."

```{r warning=F}
e_data$olre <- factor(row(e_data)[,1])
e_bin <- glmer(PAA/Control~(1|Strain)+(1|Condition)+(1|Strain:Condition)+(1|olre),
          weights=Control,
          family=binomial(link="log"),
          #control=glmerControl(optimizer="Nelder_Mead"),
          data=e_data)

# Extract BLUPs and intercept from model
e_strain_effs <- ranef(e_bin)$Strain %>% tibble::rownames_to_column(var="Strain")
e_condition_effs <- ranef(e_bin)$Condition %>% tibble::rownames_to_column(var="Condition")
e_int_effs <- ranef(e_bin)$`Strain:Condition` %>% tibble::rownames_to_column(var="Int")
e_intercept <- fixef(e_bin)[1]

# Compute BLUP of "log-reduction"
e_blups <- e_data %>%
  select(Strain,Condition) %>%
  mutate(Int=paste(Strain,Condition,sep=":")) %>%
  distinct() %>%
  left_join(e_strain_effs,by="Strain") %>% rename(strain_eff=`(Intercept)`) %>%
  left_join(e_condition_effs,by="Condition") %>% rename(condition_eff=`(Intercept)`) %>%
  left_join(e_int_effs,by="Int") %>% rename(int_eff=`(Intercept)`) %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         Int=factor(Int)) %>%
  mutate(intercept=e_intercept) %>%
  mutate(BLUP=log10(exp(1))*(-intercept-strain_eff-condition_eff-int_eff))

# Add zeroes for plotting
e_blups <- e_data %>% 
  group_by(Strain,Condition) %>%
  summarize(zeroes=sum(PAA==0)) %>%
  left_join(e_blups,by=c("Strain","Condition")) %>%
  ungroup()
```

Giving this model:
```{r}
summary(e_bin)
```

With the following CIs for variance components:
```{r warning=F}
confint(e_bin,method="boot",nsim=100) %>%
  kable()
```

This gives us the following BLUPs:
```{r}
e_strain_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()

e_condition_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()

e_int_effs %>%
  mutate(`(Intercept)`=log10(exp(1))*`(Intercept)`) %>%
  kable()
```

## Plotting the results

```{r}
e_blups %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,BLUP,zeroes) %>%
ggplot(aes(x=i_Strain,y=BLUP,color=Condition,linetype=Condition)) +
  geom_line() +
  geom_text(aes(label=zeroes)) +
  ylim(0,9) +
  labs(x="Strain",y="Binomial proportion model",title="E. coli") +
  scale_x_continuous(breaks=1:length(levels(factor(e_blups$Strain))),
                     labels=levels(factor(e_blups$Strain))) +
  theme_bw()
```

Compare to the mean estimates (with detection limit substituted where sanitizer
count is zero)

```{r}
e_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Strain,y=mean,color=Condition,linetype=Condition)) +
  geom_line() +
  geom_point() +
  labs(x="Strain",y="Mean",title="E. coli") +
  ylim(0,9) +
  scale_x_continuous(breaks=1:length(levels(factor(e_data$Strain))),
                     labels=levels(factor(e_data$Strain))) +
  theme_bw()
```

Looking at this the other way:

```{r}
e_blups %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,BLUP,zeroes) %>%
ggplot(aes(x=i_Condition,y=BLUP,color=Strain,linetype=Strain)) +
  geom_line() +
  geom_text(aes(label=zeroes)) +
  ylim(0,9) +
  labs(x="Condition",y="Binomial proportion model",title="E. coli") +
  scale_x_continuous(breaks=1:length(levels(factor(e_blups$Condition))),
                     labels=levels(factor(e_blups$Condition))) +
  theme_bw()
```

Compare to the mean estimates (with detection limit substituted where sanitizer
count is zero)

```{r}
e_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Condition,y=mean,linetype=Strain, pch=Strain)) +
  geom_line(size=1) +
geom_point(size=2.5) +
  labs(x="Condition",y="Log reduction CFU/mL",title=expression(italic("E. coli"))) +
  ylim(0,9) +
  scale_x_continuous(breaks=1:length(levels(factor(e_data$Condition))),
              labels=levels(factor(e_data$Condition))) +
  theme_bw(base_size = 20)
```

Lets try making a bar graph

```{r}
e_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  group_by(Strain,Condition) %>%
  summarize(mean=mean(reduc)) %>%
  ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,mean) %>%
ggplot(aes(x=i_Condition,y=mean, fill=factor(Strain))) +
geom_bar(stat="identity", position = "dodge")+
  labs(x="Condition",y="Log reduction CFU/mL",title=expression(italic("E. coli"))) +
  ylim(0,9) +
   scale_fill_discrete(name="Strain",
                      breaks=c(1, 2,3,4,5,6),
                      labels=c("FSL R9-5258", "FSL R9-5271", "FSL R9-5515", "FSL R9-55516", "FSL R9-5517", "TVS 353"))+
  scale_x_continuous(breaks=1:length(levels(factor(e_data$Condition))),
    labels=levels(factor(e_data$Condition))) +
  theme_bw(base_size = 15)

##another graph with jitter?

e_data_w_surrogate %>%
  mutate(reduc=ifelse(is.infinite(reduc),right_censor_limit_reduc,reduc)) %>%
  #group_by(Strain,Condition) %>%
  #summarize(mean=mean(reduc)) %>%
  #ungroup() %>%
  mutate(Strain=factor(Strain),
         Condition=factor(Condition),
         i_Strain=as.integer(Strain),
         i_Condition=as.integer(Condition)) %>%
  select(Strain,Condition,i_Strain,i_Condition,reduc) %>%
ggplot(aes(x=i_Condition,y=reduc, color=Strain)) +
geom_jitter()+
  labs(x="Condition",y="Log reduction CFU/mL",title=expression(italic("E. coli"))) +
  ylim(0,9) +
  scale_x_continuous(breaks=1:length(levels(factor(e_data$Condition))),
              labels=levels(factor(e_data$Condition))) +
  theme_bw(base_size = 20)


```


```{r}

## Prediction intervals

```{r}
mu <- e_bin@beta[1]
sds <- e_bin@theta
e_lin_pred_overall <- function(x) {
  -log10(exp(mu + 
        rnorm(1,0,sds[1]) +
        rnorm(1,0,sds[2]) +
        rnorm(1,0,sds[3]) +
        rnorm(1,0,sds[4])))
}
e_overale_sim <- data.frame(fixed_feature="None",
           fixed="None",
           sim=unlist(lapply(1:1e4,FUN=e_lin_pred_overall)))

e_lin_pred_strain_fixed <- function(x,strain) {
  -log10(exp(mu + 
        rnorm(1,0,sds[1]) +
        rnorm(1,0,sds[2]) +
        rnorm(1,0,sds[3]) +
        strain))
}
e_strain_sim <- data.frame(fixed_feature=NULL,
                     fixed=NULL,
                     sim=NULL)
for( s in e_strain_effs$Strain ) {
  tmp <- data.frame(fixed_feature="Strain",
                    fixed=s,
                    sim=unlist(
                      lapply(1:1e4,
                             FUN=e_lin_pred_strain_fixed,
                             strain=e_strain_effs$`(Intercept)`[e_strain_effs$Strain==s])))
  e_strain_sim <- rbind(e_strain_sim,tmp)
}

e_lin_pred_condition_fixed <- function(x,condition) {
  -log10(exp(mu + 
               rnorm(1,0,sds[1]) +
               rnorm(1,0,sds[2]) +
               condition +
               rnorm(1,0,sds[4])))
}
e_condition_sim <- data.frame(fixed_feature=NULL,
                     fixed=NULL,
                     sim=NULL)
for( s in e_condition_effs$Condition ) {
  tmp <- data.frame(fixed_feature="Condition",
                    fixed=s,
                    sim=unlist(
                      lapply(1:1e4,
                             FUN=e_lin_pred_condition_fixed,
                             condition=e_condition_effs$`(Intercept)`[e_condition_effs$Condition==s])))
  e_condition_sim <- rbind(e_condition_sim,tmp)
}

e_pred <- rbind(e_overale_sim,
                e_strain_sim,
                e_condition_sim)
e_pred %>%
  group_by(fixed,fixed_feature) %>%
  summarize(mean=mean(sim),
            lower=quantile(sim,0.025),
            upper=quantile(sim,0.975)) %>%
ggplot(aes(x=fixed,y=mean)) +
  geom_point() +
labs(x="fixed effects",y="Log reduction CFU/mL",title="STEC") +
  geom_errorbar(aes(ymin=lower,ymax=upper),width=0.2) +
  facet_grid(~fixed_feature,scales = "free") +
  theme(axis.text.x=element_text(angle = 45,hjust=1))
```

